Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-03-23 ms_pre_post.Rmd
Fix reference to supporting information

Despite the paucity of theory and resulting predictions, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma20134; @Arct2015; but see @Reid2015a]. However, relatively few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in diverse species, spanning sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and a tropical butterfly [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature alongside polyandry. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and polyandry should be favoured by selection when all combinations of phenotypes can evolve. Consequently, there is no theory that allows the diversity of existing empirical results regarding pre-copulatory and post-copulatory inbreeding avoidance to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@@

Post-copulatory inbreeding avoidance cannot be realised if females are not polyandrous, and is most effective for highly polyandrous females that can choose amongst multiple mates and therefore bias fertilisation accordingly. In contrast, pre-copulatory inbreeding avoidance mechanisms are most critical for females that mate only once and therefore have no post-copulatory opportunity to avoiding inbreeding. This theory is borne out in our simulation results, as selection for, and consequent evolution of, post-copulatory inbreeding avoidance was negligible in populations where polyandry did not evolve, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). Evolution of pre-copulatory inbreeding avoidance was also typically slightly stronger when polyandry did not evolve (e.g., Figure 3A vs 3B; see also Supporting Information p. S1-1). In addition to polyandry causing evolution of post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

>>>>

Despite the paucity of theory and resulting predictions, numerous empirical studies on diverse species have tested for, and in some cases found evidence of, female inbreeding avoidance in systems where polyandry also occurs [@Tregenza2002; @Varian-Ramos2012; @Kingma2013; @Arct2015; but see @Reid2015a]. However, relatively few studies have determined whether inbreeding avoidance is enacted through pre-copulatory or post-copulatory mechanisms. Pre-copulatory inbreeding avoidance has been reported in diverse species, spanning sweet potato weevils [*Cylas formicarius*, @Kuriwada2011], purple-crowned fairy-wrens [*Malurus coronatus*, @Kuriwada2011], and a tropical butterfly [*Bicyclus anynana*, @Fischer2015], while post-copulatory inbreeding avoidance has been reported in red junglefowl [*Gallus gallus*, @Pizzari2004] and crickets [*Teleogryllus oceanicus*, @Simmons2006; *Gryllus bimaculatus*, @Bretman2009]. Evidence for both pre-copulatory and post-copulatory inbreeding avoidance is available across different studies of Trinidadian guppies [*Poecilia reticulata*; @Gasparini2011; @Daniel2015] and house mice [*Mus domesticus*; @Potts1991; @Firman2015]. Meanwhile, @Liu2014 found evidence of pre-copulatory but not post-copulatory inbreeding avoidance within a single study of cabbage beetles (*Colaphellus bowringi*). However, @Ala-Honkola2011 and @Tan2012 found no evidence for pre-copulatory or post-copulatory inbreeding avoidance in fruit flies (*Drosophila melanogaster*), respectively, and Reid et al. [-@Reid2014; -@Reid2015] found no net inbreeding avoidance in song sparrows (*Melospiza melodia*) despite strong inbreeding depression. Taken together, these studies demonstrate that diverse combinations of pre-copulatory and post-copulatory inbreeding avoidance occur in nature alongside polyandry. However, there is as yet no theory that predicts what combinations of pre-copulatory and post-copulatory inbreeding avoidance and polyandry should be favoured by selection when all combinations of phenotypes can evolve. Consequently, there is no theory that allows the diversity of existing empirical results regarding pre-copulatory and post-copulatory inbreeding avoidance to be interpreted, and no clear hypotheses that could be tested through future empirical studies of individual systems or subsequent comparative analyses.

@@

Post-copulatory inbreeding avoidance cannot be realised if females are not polyandrous, and is most effective for highly polyandrous females that can choose amongst multiple mates and therefore bias fertilisation accordingly. In contrast, pre-copulatory inbreeding avoidance mechanisms are most critical for females that mate only once and therefore have no post-copulatory opportunity to avoiding inbreeding. This theory is borne out in our simulation results, as selection for, and consequent evolution of, post-copulatory inbreeding avoidance was negligible in populations where polyandry did not evolve, resulting in high drift of allele values over generations due to the inability of females to express post-copulatory choice (e.g., Figure 4C,D). Evolution of pre-copulatory inbreeding avoidance was also typically slightly stronger when polyandry did not evolve (e.g., Figure 3A vs 3B; see also Supporting Information p. S1). In addition to polyandry causing evolution of post-copulatory inbreeding avoidance, polyandry might also covary positively with post-copulatory inbreeding avoidance due to the feed back that post-copulatory inbreeding avoidance has on facilitating evolution of polyandry itself, as observed in our model (Figure 1). It would therefore be interesting to test the hypothesis that, across taxa, occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandry. To test this hypothesis, more work is needed to identify the degree to which females of different species engage in polyandry and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.
